Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2020-09-11 Paper/Manuscript_Stats&Comp/manuscript_ST&CO.tex
Minor edits to new paragraph on the diagnostic

In order to assess the performance of the models as a function of the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model, and different number of knots for the spline model, have been fitted. Figure \ref{fig12_MSE_exI_inter} shows the standardized root mean squared error (SRMSE) for interpolation and extrapolating data as a function of the number of basis functions and knots. The SRMSE is computed against the data-generating model. From Figures \ref{fig11_Posteriors_exI} and \ref{fig12_MSE_exI_inter}, it is seen that the HSGP method yield a good approximation of the exact GP model for both interpolation and extrapolation. However, the spline model does not extrapolate data properly. Both models show roughly similar interpolating performance.

In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended. 
%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern}. Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$. 
Figure \ref{fig13_diagnosis} depicts the diagnosis applied for this case study. In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnosis}. In the second iteration, $m$ is increased up to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig7_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$. As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{fig12_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP). In the third iteration, $m$ is increased up to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships brings a minimum approachable length-scale of $\ell^*=0.15$. As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{fig12_MSE_exI_inter}-left). In the fourth iteration, $m$ is increased up to $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum approachable length-scale generated by the model of relationships is $\ell^*=0.11$, and as $\hat{\ell} > \ell^*$ then approximation is diagnosed to be accurate.

>>>>

In order to assess the performance of the models as a function of the number of basis functions and number of knots, different models with different number of basis functions for the HSGP model, and different number of knots for the spline model, have been fitted. Figure \ref{fig12_MSE_exI_inter} shows the standardized root mean squared error (SRMSE) for interpolation and extrapolating data as a function of the number of basis functions and knots. The SRMSE is computed against the data-generating model. From Figures \ref{fig11_Posteriors_exI} and \ref{fig12_MSE_exI_inter}, it is seen that the HSGP method yields a good approximation of the exact GP model for both interpolation and extrapolation. However, the spline model does not extrapolate data properly. Both models show roughly similar interpolating performance.

%In practice, as the high computational demands of exact GPs prevents their use to assess the performance of the approximation, the diagnosis of the approximation as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency} is recommended.

%The diagnosis consists in comparing the estimated length-scale $\hat{\ell}$ for a given $m$ and $c$, with that minimum length-scale $\ell^*$ that can be accurately approximated, for those given $m$ and $c$, generated by the model of relationships in Figures \ref{fig6_lscale_vs_J_vs_c} and \ref{fig7_lscale_vs_J_vs_c_Matern}. Thus, if $\hat{\ell} < \ell^*$, $m$ needs to increased or $c$ decreased, and this can be iteratively until we get $\hat{\ell} \geq \ell^*$. 
Next, we demonstrate how to apply the diagnostic for approximation quality as a function of $m$ and $c$ as proposed in Section \ref{subsec_dependency}.
Figure \ref{fig13_diagnosis} depicts the diagnosis applied to case study. In the first iteration, we choose $m=20$ and $c=1.5$, which turns out to be an invalid combination of $m$ and $c$, as $c=1.5$ is too small for $m=20$ or $m=20$ is too small for $c=1.5$, as can be seen in the corresponding plot of Figure \ref{fig13_diagnosis}. In the second iteration, $m$ is increased to $m=30$, the HSGP estimates a length-scale of $\hat{\ell}=0.12$, and the minimum length-scale that can be accurately approximated, for those $m$ and $c$, generated by the model of relationships in Figure \ref{fig7_lscale_vs_J_vs_c_Matern} is $\ell^*=0.19$. As $\hat{\ell} < \ell^*$, the approximation is diagnosed to be inaccurate (see Figure \ref{fig12_MSE_exI_inter}-left where the interpolation error of HSGP is higher than that of GP). In the third iteration, $m$ is increased to $m=40$, the HSGP model estimates a length-scale of $\hat{\ell}=0.14$, and the model of relationships suggests a minimum length-scale of $\ell^*=0.15$. As $\hat{\ell}$ is still lower than $\ell^*$, the approximation is diagnosed to be inaccurate, but since $\hat{\ell}$ is quite close to $\ell^*$, the error is expected to be small (see errors in Figure \ref{fig12_MSE_exI_inter}-left). In the fourth iteration, $m$ is increased up $m=60$, the estimated length-scale is $\hat{\ell}=0.14$ and the minimum length-scale generated by the model of relationships is $\ell^*=0.11$, and the approximation is finally diagnosed to be accurate as $\hat{\ell} > \ell^*$.
